---
title: "Chocolate-Bar-Ratings"
date: "`r format(Sys.Date())`"
output: github_document
---   
   
      
```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(knitr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(readr)
library(maps)
```    
   
      
This is a dataset of 1,700 individual chocolate bars. This dataset has information on their regional origin, percentage of cocoa, the variety of chocolate bean used and where the beans were grown. I think it would be fun to use some R code to take a look at what chocolate bar gets the highest rating over the years, what countries has the highest rated bars and what goes into getting a highly rated chocolate bar.   

   
```{r, message=FALSE, warning=FALSE, echo=FALSE}
cocoa <- read.csv("flavors_of_cacao.csv", header = T, na.strings = (""))

```   
   
A look at the data:   
```{r}
kable(head(cocoa), format = "html")
```  
   
I need to remove the % sign in Cocoa.Percent and make it an int. 
   
```{r}
cocoa$Cocoa.Percent<-gsub("%","",cocoa$Cocoa.Percent)
cocoa$Cocoa.Percent <- as.integer(cocoa$Cocoa.Percent)
str(cocoa)
```   
   
Ok now lets see a distribution of the Ratings and percent of cocoa.   
```{r}
ggplot(cocoa, aes(x=Rating)) +
      geom_histogram(color = "black", fill = "red", binwidth = .25) +
      scale_y_continuous(breaks = seq(0,400,25)) +
      labs(x = "Ratings", title = "Cocoa Ratings", y = "Count")


```    

Not bad most in the 2.5 to 4 range. The mean is `r round(mean(cocoa$Rating), 3)` and the median is `r round(median(cocoa$Rating), 3)`









